%%%%%%
%
% $Author: Yogesh $
% $Date: 29.04.2024 $
% $Path: BA24-02-Time-Series\report\Contents\en$
% $Version: 1.0 $
% $Reviewed by:Rakesh $
%
% !TeX encoding = utf8
% !TeX root = Rename
% !TeX TXS-program:bibliography = txs:///bibtex
%
%%%%%%
\chapter{Software Bill of Material}
A Software Bill of Materials (SBOM) is essential for our project, which focuses on data versioning and air pollution prediction. The SBOM provides an inventory of all software components, libraries, and modules used in our project, ensuring transparency and traceability. By documenting the origins, licenses, and versions of each component, the SBOM helps manage software supply chain risks, ensuring our data versioning tools and prediction models are secure and compliant. This is crucial for maintaining data integrity and reproducibility in our air pollution predictions. Additionally, the SBOM aids in efficient incident response and vulnerability management, supporting best practices in our software development process.

\section{Tools}
\begin{itemize}
	\item \textbf{FOSSology}:
	\begin{itemize}
		\item FOSSology is an open source license compliance software system and toolkit.
		It allows users to scan software packages and analyze their licenses to
		ensure compliance with open source licensing requirements
		\cite{FOSSology:2017}.
		
		\item Website: \url{https://www.fossology.org/}
	\end{itemize}
	
	\item \textbf{SW360}:
	\begin{itemize}
		\item SW360 is an open source software component management system. It
		helps organizations manage their software components, dependencies, and
		licenses, providing a centralized repository for tracking and analyzing software
		assets \cite{SW360:2024}.
		
		\item Website: \url{https://www.eclipse.org/sw360/}
	\end{itemize}
	
	\item \textbf{FOSSA}:
	\begin{itemize}
		\item FOSSA is a management tool for open source components and licenses.
		It provides automated license compliance, dependency analysis, and
		vulnerability scanning for software projects, helping organizations ensure
		the legal and security compliance of their codebases \cite{FOSSA:2024}.
		
		\item Website: \url{https://fossa.com/}
	\end{itemize}
	
	\item \textbf{Nexus Repository}:
	\begin{itemize}
		\item Nexus Repository is a repository manager used for managing components
		and dependencies. It provides a centralized location for storing and
		distributing software artifacts, allowing organizations to control access,
		enforce policies, and track dependencies across projects \cite{ Sonatype:2024}.
		
		\item Website: \url{https://www.sonatype.com/nexus/repository-oss}
	\end{itemize}
	
	\item \textbf{JFrog Artifactory}:
	\begin{itemize}
		\item JFrog Artifactory is a universal binary repository manager that supports
		software packages created by any language or technology. It provides
		secure and efficient storage, management, and distribution of artifacts,
		helping organizations streamline their software development and delivery
		processes \cite{Artifactory:2024}.
		
		\item Website: \url{https://jfrog.com/artifactory/}
	\end{itemize}
	
	\item \textbf{WhiteSource Renovate}:
	\begin{itemize}
		\item WhiteSource Renovate helps to keep dependencies up-to-date automatically
		with continuous updates and security patches. It integrates with version
		control systems like GitHub and GitLab to monitor dependencies and
		create pull requests for updates \cite{Mend:2024}.
		
		\item Website:
		\url{https://www.whitesourcesoftware.com/free-developer-tools/renovate/}
	\end{itemize}
	
	\item \textbf{SourceClear (Now Snyk)}:
	\begin{itemize}
		\item SourceClear scans and monitors dependencies for security vulnerabilities.
		It provides real-time alerts and remediation guidance to help
		organizations identify and fix security issues in their software supply
		chain \cite{Synk:2024}.
		
		\item Website: \url{https://snyk.io/}
	\end{itemize}
	
	\item \textbf{GitLab Dependency Scanning}:
	\begin{itemize}
		\item GitLab's built-in Dependency Scanning feature automatically detects known
		vulnerabilities in project dependencies. It integrates with CI/CD
		pipelines to provide continuous security scanning and reporting
		\cite{Dependency:2024}.
		
		\item Website:
		\url{https://docs.gitlab.com/ee/user/application_security/dependency_scanning/}
	\end{itemize}
	
	\item \textbf{Black Duck}:
	\begin{itemize}
		\item Black Duck by Synopsys helps identify and manage open source components
		and their risks. It offers vulnerability scanning, license compliance,
		and policy enforcement capabilities to help organizations mitigate
		security and legal risks associated with open source usage
		\cite{BlackDuck:2024}.
		
		\item Website:
		\url{https://www.synopsys.com/software-integrity/security-testing/software-composition-analysis.html}
	\end{itemize}
\end{itemize}

\section{ Bill of Material}
Following table shows specification and the BOM for the software and packages used
in the project:
\begin{table}[H]
	\centering
	\caption{Software Bill of Materials}
	\begin{tabular}{|l|l|l|l|}
		\hline
		\textbf{Software/Framework/Package} & \textbf{Open Source/Licensed} & \textbf{Version} & \textbf{User} \\
		\hline
		PyCharm Professional                & Open Source                   & 2024.1.3         &               \\
		Jupyter Notebook                    & Open Source                   & 1.0.0            &               \\
		Python                              & Open Source                   & 3.9.5            &               \\
		Pandas                              & Open Source                   & 2.2.2            & Y.Gunjal      \\
		NumPy                               & Open Source                   & 1.26.0           &               \\
		Matplotlib                          & Open Source                   & 3.8.4            &               \\
		Seaborn                             & Open Source                   & 0.13.2           &               \\
		Streamlit                           & Open Source                   & 1.36.0           &               \\
		\hline
	\end{tabular}
\end{table}

\begin{itemize}
	\item \textbf{PyCharm 2024.1.3 (Professional):}
	\begin{itemize}
		\item PyCharm is an integrated development environment (IDE) designed for Python.
		It offers advanced features, making it a robust tool for coding,
		debugging, and testing Python applications \cite{Pycharm:2024}.
	\end{itemize}
	
	\item \textbf{Jupyter Notebook:}
	\begin{itemize}
		\item Jupyter Notebook is an open-source web application that allows you to
		create and share documents containing live code, equations, visualizations,
		and narrative text. It is widely used for data analysis, machine
		learning, and scientific computing \cite{Jupyter:2024}.
	\end{itemize}
	\subsection{ List of Packages}
	
	\item \textbf{Matplotlib:}
	\begin{itemize}
		\item Matplotlib is a versatile 2D plotting library for Python. It
		empowers users to create static, animated, and interactive plots, making
		it a valuable tool for visualizing data \cite{Matplotlib:2024}.
	\end{itemize}
	
	\item \textbf{Numpy:}
	\begin{itemize}
		\item Numpy is a powerful library for numerical computing in Python. It
		supports large, multi-dimensional arrays and matrices, along with a
		variety of mathematical functions, essential for scientific computing \cite{NumPy:2024}.
	\end{itemize}
	
	\item \textbf{Pandas:}
	\begin{itemize}
		\item Pandas is a robust data manipulation and analysis library for Python.
		Its key data structure, DataFrame, facilitates efficient data handling
		and manipulation tasks \cite{Pandas:2024}.
	\end{itemize}
	
	\item \textbf{Seaborn:}
	\begin{itemize}
		\item Seaborn is a powerful data visualization library built on top of Matplotlib.
		It provides a high-level interface for creating attractive and
		informative statistical graphics. Seaborn simplifies the process of visualizing
		complex datasets by offering a wide range of plotting functions and built-in
		themes. With Seaborn, users can easily create aesthetically pleasing plots
		for exploring relationships between variables, identifying patterns, and
		communicating insights effectively \cite{Seaborn:2024}.
	\end{itemize}
	
	\item \textbf{scikit-learn:}
	\begin{itemize}
		\item Scikit-learn (sklearn) is an open-source machine learning library for
		Python. It provides simple and efficient tools for data mining and data
		analysis, supporting various supervised and unsupervised learning algorithms.
		Scikit-learn is built on Numpy, SciPy, and Matplotlib, making it an essential
		library for machine learning projects \cite{Scikit-Learn:2024}.
	\end{itemize}
	
	\item \textbf{XGBoost:}
	\begin{itemize}
		\item XGBoost (Extreme Gradient Boosting) is a scalable and efficient machine
		learning library for gradient boosting. It is widely used for
		classification, regression, and ranking tasks in structured or tabular
		data. XGBoost offers optimized performance through parallel and distributed
		computing, making it a popular choice for competitive machine learning
		\cite{xgboost:2024}.
	\end{itemize}
	
	\item \textbf{Streamlit:}
	\begin{itemize}
		\item Streamlit is an open-source Python library that makes it easy to create
		and share custom web applications for machine learning and data science.
		With a few lines of code, users can build interactive dashboards and applications,
		allowing for rapid prototyping and deployment of data-driven solutions. Streamlit
		simplifies the development process by automatically updating the app as the
		code changes, facilitating real-time collaboration and exploration
		\cite{Streamlit:2024}.
	\end{itemize}
\end{itemize}